Subject-level developmental relationships

### 2024-09-16 3_developmental_relationships.qmd
add residualized age analysis and fix a couple bugs
- add residualized age correlations
- make sure trials are loaded
- move up dataset subsetting to those with age variation
- update rt_helper to most current version

```{r}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(source(here("helper","common.R")))
d_sub <- readRDS(here("cached_intermediates","1_d_sub.Rds"))
```

@@

```{r, fig.width=9, fig.height=9}
d_sub |>
  select(log_age, log_rt,  
         long_window_accuracy,  prod, comp) |>
GGally::ggpairs(progress = FALSE, 
                lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

@@

```{r}
d_sub_agevar <- d_sub |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name)

@@

@@

ggplot(filter(d_sub, 
              !is.na(rt), !is.na(factor_rt), 
              dataset_name %in% datasets_with_age_variation$dataset_name),
       aes(x = age, 
           y = long_window_accuracy, 
           col = factor_rt)) +
  geom_point(alpha = .3) +
  geom_smooth(se= FALSE, method = "lm") +
  scale_x_log10() + 
  geom_hline(yintercept = .5, lty = 2)+
  viridis::scale_color_viridis(discrete = TRUE, 
                               name = "RT quantile") +
  facet_wrap(~dataset_name) + 
  xlab("Age (log months)") + 
  ylab("Accuracy (long window)")

>>>>

```{r}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(source(here("helper","common.R")))
d_sub <- readRDS(here("cached_intermediates","1_d_sub.Rds"))
d_trial <- readRDS(here("cached_intermediates","1_d_trial.Rds"))
```

@@

```{r, fig.width=9, fig.height=9}
d_sub |>
  select(log_age, log_rt,  
         long_window_accuracy,  prod, comp) |>
GGally::ggpairs(progress = FALSE, 
                lower = list(continuous = GGally::wrap("points", alpha = 0.03)),
                axisLabels = "show")

@@

### Controlling for age (between dataset)

```{r}
d_residualized_by_age <- d_sub |>
  select(subject_id, administration_id,log_age,log_rt, long_window_accuracy, prod, comp) |> 
  pivot_longer(cols=c(log_rt,long_window_accuracy,  prod, comp),names_to="measure",values_to="value") |>
  group_by(measure) |> 
  nest() |> 
  mutate(fits = map(data, \(d) lm(value ~ log_age, data = d) |>
                      broom::augment(newdata = d))) |> 
  unnest(fits) |> 
  select(-data)

d_residualized_by_age |>
  select(administration_id, measure, .resid) |>
  pivot_wider(names_from = measure, values_from = .resid) |>
  select(-administration_id) |>
  GGally::ggpairs(progress=FALSE,lower = list(continuous = GGally::wrap("points", alpha = 0.03)))
```

@@

d_sub_agevar <- d_sub |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name)

@@

Hmm, want to think more about  estimating an "average" correlation when the variation in age changes so dramatically between datasets

```{r}
#order by correlation value of log age and long window accuracy
all_corrs_log_age_acc <- all_corrs |> 
  ungroup() |> 
  filter(variable1 == "log_age" & variable2 == "long_window_accuracy") |>
  mutate(dataset_name_ordered= fct_reorder(dataset_name,r))

d_sub_agevar <- d_sub_agevar |> 
  left_join(all_corrs_log_age_acc)

#quick peek at cross-dataset estimates, ranked by strength of correlation
ggplot(d_sub_agevar,aes(log_age,long_window_accuracy))+geom_point()+geom_smooth(method="lm")+facet_wrap(~dataset_name_ordered)
```

### Controlling for age (within dataset)

```{r}
all_corrs_age_residualized <- d_sub_agevar |>
  ungroup() |> 
  #compute variables residualized by age
select(dataset_name, administration_id,subject_id,log_age, log_rt, rt,
                       long_window_accuracy, long_window_elogit,prod) |>
  pivot_longer(cols=c(log_rt, rt,
                       long_window_accuracy, long_window_elogit,  prod),names_to="measure",values_to="value") |>
  group_by(dataset_name,measure) |> 
  nest() |> 
  mutate(fits = map(data, \(d) {if(sum(!is.na(d$value))>0) {lm(value ~ log_age, data = d) |> broom::augment(newdata = d)} })) |>
  unnest(fits) |> 
  select(dataset_name,subject_id,administration_id,measure, .resid) |> 
  rename(resid = .resid) |> 
  pivot_wider(names_from=measure,values_from=resid,names_prefix="resid_") |> 
  #now group by dataset and compute pairwise correlations
  select(-subject_id,-administration_id) |>
  group_by(dataset_name) |>
  nest() |>
  mutate(corrs = map(data, ~cor(.x, use = "pairwise.complete.obs")),
         long_corrs = map(corrs, ~pivot_longer(as_tibble(.x) |> 
                                                mutate(variable1 = colnames(.x)), 
                                               -variable1, names_to = "variable2", 
                                               values_to = "r"))) |>
  mutate(corrs = map(data, ~cor(.x, use = "pairwise.complete.obs")),
         long_corrs = map(corrs, ~pivot_longer(as_tibble(.x) |> 
                                                mutate(variable1 = colnames(.x)), 
                                               -variable1, names_to = "variable2", 
                                               values_to = "r"))) |>
  select(-data, -corrs) |>
  unnest(cols = c(long_corrs)) |>
  filter(variable1 != variable2, 
         !(variable1 == "resid_rt" & variable2 == "resid_log_rt"), 
         !(variable1 == "resid_log_rt" & variable2 == "resid_rt"),
         !(variable1 == "resid_long_window_accuracy" & variable2 == "resid_long_window_elogit"),
         !(variable1 == "resid_long_window_elogit" & variable2 == "resid_long_window_accuracy"))

ggplot(all_corrs_age_residualized,
       aes(x = variable1, y = r, col = dataset_name)) + 
  geom_point(alpha = .5, pch = 20) +
  stat_summary(fun.data = "mean_cl_boot", aes(group = 1)) +
  geom_hline(yintercept = 0, lty = 2) + 
  # coord_flip() +
  facet_grid(.~variable2, scales = "free_x") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

@@

```{r}
ggplot(filter(d_sub, 
              !is.na(rt), !is.na(factor_rt), 
              dataset_name %in% datasets_with_age_variation$dataset_name),
       aes(x = age, 
           y = long_window_accuracy, 
           col = factor_rt)) +
  geom_point(alpha = .3) +
  geom_smooth(se= FALSE, method = "lm") +
  scale_x_log10() + 
  geom_hline(yintercept = .5, lty = 2)+
  viridis::scale_color_viridis(discrete = TRUE, 
                               name = "RT quantile") +
  facet_wrap(~dataset_name) + 
  xlab("Age (log months)") + 
  ylab("Accuracy (long window)")
